---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# oecdoda

<!-- badges: start -->
![R CMD Check](https://github.com/tidy-intelligence/r-oecdoda/actions/workflows/R-CMD-check.yaml/badge.svg)
![Lint](https://github.com/tidy-intelligence/r-oecdoda/actions/workflows/lint.yaml/badge.svg)
[![Codecov test coverage](https://codecov.io/gh/tidy-intelligence/r-oecdoda/graph/badge.svg)](https://app.codecov.io/gh/tidy-intelligence/r-oecdoda)
<!-- badges: end -->

Access and Analyze Official Development Assistance (ODA) using the [OECD API](https://gitlab.algobank.oecd.org/public-documentation/dotstat-migration/-/raw/main/OECD_Data_API_documentation.pdf). ODA data includes sovereign-level aid data such as key aggregates (DAC1), geographical distributions (DAC2A), project-level data (CRS), and multilateral contributions (Multisystem).

The package is part of the [EconDataverse](https://www.econdataverse.org/) family of packages aimed at helping economists and financial professionals work with sovereign-level economic data. For a Python implementation with a similar interface, see [`oda-reader`](https://github.com/ONEcampaign/oda_reader). 

This package is a product of Christoph Scheuch and not sponsored by or affiliated with the OECD in any way, except for the use of the OECD API.

## Installation

You can install `oecdoda` from CRAN via:

``` r
install.packages("oecdoda")
```

You can install the development version of `oecdoda` from [GitHub](https://github.com/tidy-intelligence/r-oecdoda) with:

``` r
# install.packages("pak")
pak::pak("tidy-intelligence/r-oecdoda")
```

## Usage

Load the package:

```{r}
library(oecdoda)
```

`oecdoda` simplifies access to multiple datasets from the OECD API. Each dataset provides specific filters, which can be listed using:

```{r}
oda_list_filters()
```

### Downloading DAC1 Data

The DAC1 dataset includes key aggregates of ODA flows and grant equivalents from DAC members and other providers, as well as other resource flows to developing countries.

```{r}
oda_get_dac1(
  start_year = 2018,
  end_year = 2022,
  filters = list(
    donor = c("FRA", "USA"),
    measure = 11017,
    flow_type = 1160,
    unit_measure = "XDC",
    price_base = "V"
  )
)
```

### Downloading DAC2A Data

The DAC2A dataset provides data on the geographical distribution of bilateral and multilateral ODA disbursements to developing countries and territories on the DAC List of ODA recipients. Data is available by donor, recipient, and aid type (e.g., grants, loans, technical cooperation, or philanthropic flows).

```{r}
oda_get_dac2a(
  start_year = 2018,
  end_year = 2022,
  filters = list(
    donor = "GBR",
    recipient = c("GTM", "CHN"),
    measure = 106,
    price_base = "Q"
  )
)
```

### Downloading CRS Data

The CRS dataset provides granular, activity-level statistics on who provides what aid, to where, and for what purpose, on a flow basis or a grant-equivalent basis.

```{r}
oda_get_crs(
  start_year = 2018,
  end_year = 2022,
  filters = list(
    donor = c("AUT", "FRA", "USA"),
    recipient = "BIH",
    measure = 100,
    channel = 60000,
    price_base = "Q"
  )
)
```

### Downloading Multisystem Data

The Multisystem dataset presents providers’ total use of the multilateral system, including both core contributions to multilateral organizations and bilateral aid channeled through these organizations.

```{r}
oda_get_multisystem(
  start_year = 2018,
  end_year = 2022,
  filters = list(
    donor = "DAC",
    recipient = "DPGC",
    sector = 1000,
    measure = 10
  )
)
```

### Rate Limiting

`oecdoda` automatically handles the limits of the OECD API of 20 calls per minute. If you want to change the limits, use the following options: 

```{r}
options(
  oecdoda.rate_capacity = 10,
  oecdoda.rate_fill_time = 60
)
```

## Contributing

Contributions to `oecdoda` are welcome! If you'd like to contribute, please follow these steps:

1. **Create an issue**: Before making changes, create an issue describing the bug or feature you're addressing.
2. **Fork the repository**: After receiving supportive feedback from the package authors, fork the repository to your GitHub account.
3. **Create a branch**: Create a branch for your changes with a descriptive name.
4. **Make your changes**: Implement your bug fix or feature.
5. **Test your changes**: Run tests to ensure your changes don't break existing functionality.
6. **Submit a pull request**: Push your changes to your fork and submit a pull request to the main repository.
